---
title: "hw02"
author: "Kaila An"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-libraries}
library(tidyverse)
library(data.table)
library(dplyr)
library(dtplyr)
library(ggplot2)
```

## Data Wrangling

Read in the data.

```{r read}
chsind <- data.table::fread("chs_individual.csv")
chsreg <- data.table::fread("chs_regional.csv")
```

Merge the data sets.

```{r merge}
chs <- merge(chsind, chsreg, by="townname")
```

#1. After merging the data, make sure you don’t have any duplicates by counting the number of rows. Make sure it matches.

```{r}
nrow(chsind)
nrow(chs)
```

There is no repeat data.  The merged set and the original individual set both have 1200 rows.

#2. Create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24). To make sure the variable is rightly coded, create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category.

```{r obesity_level}
chs$obesity_level <- as.factor (ifelse(chs$bmi < 14, 'underweight',
                                ifelse(chs$bmi < 22, 'normal', 
                                ifelse(chs$bmi < 24, 'overweight','obese' ))))
```

```{r summary table}
summary <- chs[, .(
  minimum = min(bmi),
  maximum = max(bmi),
  count   = n
)]
```

#3. Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove.” The variable should have four categories in total.

```{r smoke gas}
chs[ , smoke_gas_exposure := fifelse(smoke == 0 & gasstove == 0, "Neither",
                            fifelse(smoke == 1 & gasstove == 0, "smoke_only",
                            fifelse(smoke == 0 & gasstove == 1, "gas_only",
                            fifelse(smoke == 1 & gasstove == 1, "both", NA_character_))))]

chs[ ,.(total = length(smoke)), by = smoke_gas_exposure]
```


#4. Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”




## Looking at the Data (EDA)


Primary Questions of Interest

#1. What is the association between BMI and FEV (forced expiratory volume)?

#2. What is the association between smoke and gas exposure and FEV?

#3. What is the association between PM2.5 exposure and FEV?


Follow the EDA checklist from week 3 and the previous assignment. Be sure to focus on the key variables. Visualization Create the following figures and interpret them. Be sure to include easily understandable axes, titles, and legends.


#1. Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.

#2. Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.

#3. Barchart of BMI by smoke/gas exposure.

#4. Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.

#5. A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.

#6. Choose a visualization to examine whether PM2.5 mass is associated with FEV.


