---
title: "hw02"
author: "Kaila An"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-libraries}
library(tidyverse)
library(data.table)
library(dplyr)
library(dtplyr)
library(ggplot2)
```

## Data Wrangling

Read in the data.

```{r read}
chsind <- data.table::fread("chs_individual.csv")
chsreg <- data.table::fread("chs_regional.csv")
```

Merge the data sets.

```{r merge}
chs <- merge(chsind, chsreg, by="townname")
```


#1. After merging the data, make sure you don’t have any duplicates by counting the number of rows. Make sure it matches.

```{r}
nrow(chsind)
nrow(chs)
```

There is no repeat data.  The merged set and the original individual set both have 1200 rows.


#2. Create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24). To make sure the variable is rightly coded, create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category.

```{r obesity_level}
chs$obesity_level <- as.factor (ifelse(chs$bmi < 14, 'underweight',
                                ifelse(chs$bmi < 22, 'normal', 
                                ifelse(chs$bmi < 24, 'overweight','obese' ))))
```

```{r impute}
chs[ , bmi_imp := fcoalesce(bmi, mean(bmi, na.rm = TRUE)),
by = . (hispanic, male)]
```

```{r summary table}
chs[, .(
  minimum = min(bmi),
  maximum = max(bmi),
  count   = .N
), by = c("obesity_level")]
```


#3. Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove.” The variable should have four categories in total.

```{r smoke gas}
chs[ , smoke_gas_exposure := ifelse(smoke == 0 & gasstove == 0, "Neither",
                             ifelse(smoke == 1 & gasstove == 0, "smoke_only",
                             ifelse(smoke == 0 & gasstove == 1, "gas_only",
                             ifelse(smoke == 1 & gasstove == 1, "both", NA_character_))))]

chs[ ,.(total = length(smoke)), by = smoke_gas_exposure]
```


#4. Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”

```{r town}
chs[,.(
   mean = mean(fev, na.rm=TRUE),
   sd   = sd(fev, na.rm=TRUE)
), by   = townname][order(townname)]
```

```{r sex}
chs[,.(
   mean = mean(fev, na.rm=TRUE),
   sd   = sd(fev, na.rm=TRUE)
), by   = male][order(male)]
```

```{r obesity level}
chs[,.(
   mean = mean(fev, na.rm=TRUE),
   sd   = sd(fev, na.rm=TRUE)
), by   = obesity_level][order(obesity_level)]
```
 
```{r smoke_gas_exposure}
chs[,.(
   mean = mean(fev, na.rm=TRUE),
   sd   = sd(fev, na.rm=TRUE)
), by   = smoke_gas_exposure][order(smoke_gas_exposure)]
```


## Looking at the Data (EDA)


Follow the EDA checklist from week 3 and the previous assignment. Be sure to focus on the key variables. Visualization Create the following figures and interpret them. Be sure to include easily understandable axes, titles, and legends.


#1. Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.

```{r facet plot}
chs[!is.na(townname)]%>%
  ggplot(data = chs, mapping = aes(x = bmi, y = fev, color = townname)) +
  geom_point() +
  facet_wrap(~townname, nrow = 3) +
  geom_smooth(method = 'lm', formula = y~x)
```


#2. Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.

```{r FEV by BMI}
ggplot(data = chs) + 
  geom_histogram(mapping = aes(x = fev, fill = obesity_level), binwidth = 50) + scale_fill_brewer(palette = "RdPu")
```

```{r FEV by smoke/gas}
ggplot(data = chs) + 
  geom_histogram(mapping = aes(x = fev, fill = smoke_gas_exposure), binwidth = 50) + scale_fill_brewer(palette = "OrRd")
```


#3. Barchart of BMI by smoke/gas exposure.

```{r BMI}
chs[!is.na(smoke_gas_exposure)] %>%
  ggplot() + 
  geom_bar(mapping = aes(x = bmi, colour = smoke_gas_exposure, fill = smoke_gas_exposure, binwidth = 100)) + 
  scale_fill_brewer(palette = "BuPu")
```

```{r bmi by obesity_levels}
chs[!is.na(smoke_gas_exposure)] %>%
  ggplot() + 
  geom_bar(mapping = aes(x = obesity_level, colour = smoke_gas_exposure, fill = smoke_gas_exposure, binwidth = 50)) + 
  scale_fill_brewer(palette = "BuPu")
```


#4. Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.

```{r fev by bmi line plot}
ggplot(data = chs) +
 geom_line(mapping = aes(x = bmi, y = fev))
```

```{r fev by obesity_level line plot}
ggplot(data = chs) +
 geom_line(mapping = aes(x = obesity_level, y = fev))
```


```{r fev by smoke/gas exposure line plot}
ggplot(data = chs) +
 geom_line(mapping=aes(x = smoke_gas_exposure, y = fev))
```
```{r fev by bmi boxplot}
chs[!is.na(bmi)] %>% 
  ggplot()+
  geom_boxplot(mapping=aes(x=bmi, y=fev, fill=bmi))
```

```{r fev by obesity_level boxplot}
chs[!is.na(obesity_level)] %>% 
  ggplot() +
  geom_boxplot(mapping = aes(x = obesity_level, y = fev, fill = obesity_level)) +
  scale_fill_brewer(palette = "Blues")
```

```{r fev by smoke/gas exposure boxplot}
chs[!is.na(bmi)] %>% 
  ggplot() +
  geom_boxplot(mapping = aes(x = smoke_gas_exposure, y = fev, fill = smoke_gas_exposure)) +   scale_fill_brewer(palette = "Blues")
```


#5. A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.

```{r leaflet}
library(leaflet)
tem.pal <- colorFactor(topo.colors(2), domain = chs$pm25_mass)
leaflet(chs) %>% 
  # the looks of the map
  addTiles() %>%
  # some circles
  addCircles (color = ~tem.pal(pm25_mass),
               label = ~pm25_mass, 
               opacity=0.01, fillOpacity = 0.01, radius = 500) %>%
  # add a pretty legend
  addLegend('bottomleft', pal= tem.pal, values = chs$pm25_mass,
             title ='pm25_mass levels in CHS communities', opacity=1)
```

#6. Choose a visualization to examine whether PM2.5 mass is associated with FEV.

```{r}
chs[!is.na(townname)]%>%
  ggplot(data = chs, mapping = aes(x = pm25_mass, y = fev)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = y~x)
```


## Primary Questions of Interest

#1. What is the association between BMI and FEV (forced expiratory volume)?

#2. What is the association between smoke and gas exposure and FEV?

#3. What is the association between PM2.5 exposure and FEV?


