---
title: "hw03"
author: "Kaila An"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-libraries}
library(dtplyr)
library(forcats)
library(rvest)
library(httr)
library(ggplot2)
library(tidyverse)
library(tidytext)
library(stringr)
```

# APIs

## Using the NCBI API, look for papers that show up under the term “sars-cov-2 trial vaccine.” Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

```{r}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")

# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/div[1]")

# Turning it into text
counts <- as.character(counts)

# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")
```

I found 4,006 papers in the database.

Details:

```{r}
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db      = "pubmed",
    term    = "sars-cov-2 trial vaccine",
    retmax  = 250
  ), 
)
# Extracting the content of the response of GET
ids <- httr::content(query_ids)
```

```{r}
# Turn the result into a character vector
ids <- as.character(ids)
# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[[:digit:]]+</Id>")[[1]]
# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "</?Id>")
head(ids)
```


## Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.

```{r}
publications <- GET(
  url        = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query      = list(
  db         = "pubmed",
  id         = paste(ids[250],collapse = ","),
  retmax     = 250,
  rettype    = "abstract"
    )
)
# Turning the output into character vector
publications     <- httr::content(publications)
publications_txt  <- as.character(publications)
```


## As we did in lab 7. Create a dataset containing the following:

Pubmed ID number,
Title of the paper,
Name of the journal where it was published,
Publication date, and
Abstract of the paper (if any).

```{r}
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
```

```{r}
abstracts <- str_extract(pub_char_list, "<Abstract>[[:print:][:space:]]+</Abstract>")
abstracts[[1]]
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]- =\"]+>")
abstracts[[1]]
abstracts <- str_replace_all(abstracts, "[[:space:]]+", " ")
abstracts[[1]]
```

```{r}
titles <- str_extract(pub_char_list, "<ArticleTitle>[[:print:][:space:]]+</ArticleTitle>")
titles[[1]]
titles <- str_remove_all(titles, "</?[[:alnum:]- =\"]+>")
titles[[1]]
```

```{r}
journal <- str_extract(pub_char_list, "<Journal>[[:print:][:space:]]+</Journal>")
journal[[1]]
journal <- str_extract(journal, "<Title>[[:print:][:space:]]+</Title>")
journal[[1]]
journal <- str_remove_all(journal, "</?[[:alnum:]- =\"]+>") 
journal[[1]]
journal <- str_replace_all(journal, "[[:space:]]+", " ")
journal[[1]]
```

```{r}
pubdate <- str_extract(pub_char_list, "<PubDate>[[:print:][:space:]]+</PubDate>")
pubdate[[1]]
pubdate <- str_remove_all(pubdate, "</?[[:alnum:]- =\"]+>")
pubdate[[1]]
pubdate <- str_replace_all(pubdate, "[[:space:]]+", " ")
pubdate[[1]]
```


```{r}
database <- data.frame(
  PubMedId = ids,
  Titles    = titles,
  Abstracts = abstracts,
  Journal = journal,
  PubDate = pubdate
)

knitr::kable(database[1:20,])
```


# Text Mining
A new dataset has been added to the data science data repository https://github.com/USCbiostats/data-science-data/tree/master/03_pubmed. The dataset contains 3241 abstracts from articles across 5 search terms. Your job is to analyse these abstracts to find interesting insights.

## Question 1. Tokenize the abstracts and count the number of each token. Do you see anything interesting? Does removing stop words change what tokens appear as the most frequent? What are the 5 most common tokens for each search term after removing stopwords?

Read in data:

```{r}
pubmed <- read.csv("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/03_pubmed/pubmed.csv")

str(pubmed)
```

```{r}
pubmed <- as_tibble(pubmed)
pubmed
```

Tokenize the abstracts:

```{r}
pubmed %>%
  unnest_tokens(token, abstract) %>%
  count(token, sort = TRUE) %>%
  top_n(20,n) %>%
  ggplot(aes(n, fct_reorder(token, n))) +
  geom_col()
```

After initially tokenizing, we do not see a ton of medical terminology.  The most common words found were 'the', 'of', 'and', 'in', etc.  These are words that are common in writings, in general.  


Remove the stop words:

```{r}
pubmed %>%
  unnest_tokens(token, abstract) %>%
  anti_join(stop_words, by = c("token" = "word")) %>%
  count(token, sort = TRUE) %>%
  filter(!grepl(pattern = "^[0-9]+$", x = token)) %>%
  top_n(5, n) %>%
  ggplot(aes(n, fct_reorder(token, n)))+
  geom_col()
```

After removing the stop words, we see more medical terminology as the most common tokens/terms in the abstracts (covid, patients, cancer, prostate, disease).


## Question 2. Tokenize the abstracts into bigrams. Find the 10 most common bigram and visualize them with ggplot2.

```{r}
pubmed %>%
  unnest_ngrams(bigram, abstract, n=2) %>%
  count(bigram, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(n, fct_reorder(bigram, n))) +
  geom_col()
```


## Question 3. Calculate the TF-IDF value for each word-search term combination. (here you want the search term to be the “document”) What are the 5 tokens from each search term with the highest TF-IDF value? How are the results different from the answers you got in question 1?

```{r}
pubmed %>%
  unnest_tokens(token, abstract) %>%
  count(term, token, sort = TRUE) %>%
  bind_tf_idf(token, term, n) %>%
  top_n(50, tf_idf) %>%
  arrange(desc(tf_idf)) %>%
  knitr::kable()
```

COVID: 
    1. Covid
    2. Pandemic
    3. Coronavirus
    4. Sars
    5. Cov
    
Prostate Cancer: 
    1. Prostate
    2. Androgen
    3. Psa
    4. Prostatectomy
    5. Castration
    
Meningitis: 
    1. Meningitis
    2. Meningeal
    3. Pachymeningitis
    4. Csf
    5. Meninges
    
Preeclampsia: 
    1. Eclampsia
    2. Preeclampsia
    3. Pregnancy
    4. Maternal
    5. Gestational
    
Cystic Fibrosis: 
    1. CF
    2. Fibrosis
    3. Cystic
    4. Cftr
    5. Sweat

The answers are different than the results received in Question #1, in that we did not have to worry about the stop words at all, and we focused strictly on medical terminology with the word-search terms.  In addition, the tokens we worked with in this question were COVID, Meningitis, Prostate Cancer, Preeclampsia, and Cystic Fibrosis, while in Question #1, the top 5 tokens were COVID, patients, cancer, prostate, and disease.  The top tokens in the two questions were similar, yet quite different.  Prostate(cancer) and COVID seem to persist as top tokens, overall.














